---
title: "ESm_206_lab5_rsr"
author: "Ruben Sanchez Ramirez"
date: "10/29/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with 'lubraidate'
- Get counts of observetions with count ()
- nNcount () rows
- One and two-sample t-test
Create a heatmap woth geom_tile()

```{r, include=FALSE}
# Read in data and attch packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"), 
                              na = "-99999") %>% 
  janitor::clean_names()
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

#### Exploratory Data Visualization

We're only going to consider 'site' as out variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
ggplot(lobster_tidy, aes(x = size_mm,)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scale = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
              

```

#### Covert the 'data' column to class 'Date"

We'll use the 'libridate' package to convert to Date format, and then to help us easily parse month & yeat

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Now let's parse year and month using lubridat::month() and lubridate::year()

```{r}

lobster_parse_data <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE), 
    obs_year = lubridate::year(date_new)
  )

```

Now let's find counts of observed lobsters based on diffrent hierarchical groupungs:

First: let's count lobsters by year and month:
```{r}
lobster_ym <- lobster_parse_data %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <- lobster_parse_data %>% 
  dplyr::count(obs_year)

lobster_y

lobster_site <- lobster_parse_data %>% 
  dplyr::count(site)

lobster_site
```
If we want to create a summary table that contains statistics OTHER then counts by group, it's easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_data %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary
```

#### Find Confidence Intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-test to compare means of tow samples (... this will be covered conceptually in lecture week 6)

```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
```

#### Two-sample t-test to compare means

We want to ask: Is there a significant diffrence in lobster lenghts at Naples and Mohawk reefs?

We've done our nessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}

naple_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, naple_sample)
mn_ttest

```

There is a significant difference in lobster lenghts between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha = 0.05).

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```

 **DONT DO:** There is a sig diff (t(1858.8) = -19.849, p = 2.2e-26).
 
 Here is the t-statistic: `r mn_ttest2$p.value`

Here is my t-statistic: `r mn_ttest2$statistic`

#### Now: a heatmap

```{r}

lobster_ys <- lobster_parse_data %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

